\chapter{Introduction} 

Discrete event simulation (DES) is a computational methodology for
modeling and analysis of a wide spectrum of systems.  In DES, the
system being modeled is logically subdivided into small, independent
but interacting subsystems with their own independent states.  The
model and implementation of a subsystem is called a Logical Process
(LP) which manage the state assocated with them.  Accordingly, a DES
is essentially designed as a set of logical processes (LPs) that
interact with each other.  LPs interact by exchanging and processing
discrete-timestamped events or messages~\cite{jafer-13}.  Processing
an event essentially introduces a change in an LP's state and causes
the LP to generate additional events (to itself or other LPs in the
model).

A key aspect of DES, is that state changes occur at discrete
times~\cite{fishman-13}.  At each point in time in a simulation, a
virtual time-stamp is assigned to an event and the event precipitates
a transition from one state to another state. This change in system
state is used to represent the dynamic nature and behavior of a
real-world system~\cite{fujimoto-90}.

DES has been used in a variety of fields in academia, industry and the
public sector as a tool to help inform our knowledge of discrete event
systems and to improve decision-making
processes~\cite{fishman-13}. DES provides an effective means for
analyzing real or artificial systems without the constraint of limited
resources such as time, financial costs, or safety. For example, the
simulation of a battlefield environment can deliver insightful
information to military planners on enemy troop movements, tactics and
capabilities during strategic planning efforts~\cite{hill-01}. A
discrete event simulation of the battlefield allows military leaders
to examine the impacts of decisions without the real-world risks
associated with committing forces to dangerous environments.

\section{Parallel Simulation}

Parallelism in computing frameworks that support DES increase
performance throughput that is needed to construct and execute large
scale and complex simulation models. With the growth and prevalence of
semiconductor technology, cheaper and powerful multi-processors can be
instrumented to achieve greater computing power for parallel discrete
event simulations (PDES).  In parallel simulations, LPs are subdivided
or partitioned to operate on different compute units.  However, event
processing on the different parallel must be synchronized to ensure
causally consistent event processing.  Consequently, the speedup
achieved using multi-core and multi-processor systems requires
efficient parallel simulations by minimizing synchronization costs.
Currently, two broad types of synchronization methods are used in
PDES, namely: conservative and optimistic approaches~\cite{jafer-13}.
Conservative methods tightly coordinate event processing so that
causal violations do not occur.  Optimistic methods, such as Time
Warp~\cite{jafer-13}, loosely synchronize LPs -- they permit temporary
causal violations to occur but detect and recover from causal
violations.  Recently, optimistic synchronization method have shown to
outperform conservative methods for certain classes of
systems~\cite{jafer-13}.


\section{Managing Pending Events}

Sequential and parallel DES are designed as a set of logical processes
(LPs) that interact with each other by exchanging and processing
timestamped events or messages~\cite{jafer-13}. Events that are yet to
be processed are called "pending events". Pending events must be
processed by LPs in priority order to maintain causality, with event
priorities being determined by their timestamps. Consequently, data
structures for managing and prioritizing pending events play a
critical role in ensuring efficient sequential and parallel
simulations~\cite{jones-86,ronngren-97,brown-88,franceschini-15}. The
effectiveness of data structures for event management is a conspicuous
issue in larger simulations, where thousands or millions of events can
be pending~\cite{carothers-2010,yeom-14}. Large pending event sets can
arise when a model has many LPs or when each LP generates / processes
many events. Overheads in managing pending events is magnified in fine
grained simulations where the time taken to process an event is very
short. Furthermore, the synchronization strategy used in PDES called,
Time Warp can further impact the effectiveness of the data structure
due to additional processing required during rollback-based recovery
operations.

\section*{Thesis Statement}

Many investigations have explored the effectiveness of a wide variety
of data structures for managing the pending event set. Among the
various data structures, the Ladder Queue proposed by
Tang,~\cite{tang-05} has shown to be the most effective data structure
for managing pending events~\cite{dickman-13,franceschini-15},
particularly in sequential DES. Our research proposes and explorers
multi-tier data structures for the improved management of the pending
event set in sequential and optimistic parallel simulations. The
objective of the research is to develop a data structure that
outperforms all other data structures in managing pending events. This
task is accomplished by comparing the effectiveness of previous and
newly defined data structures against our fine-tuned version of the
Ladder Queue~\cite{tang-05} because it has shown to be very efficient
for sequential DES. \newline

\textbf{Thesis:} multi-tiered data structures, especially the novel,
\textbf{2tLadderQ} and \textbf{3tHeap} pending event structures
outperform all other data structures in sequential and optimistic
parallel simulations.

\section*{Related Work}

Recently, Franceschini~\cite{franceschini-15} compared several
priority-queue based pending event data structures to evaluate their
performance in the context of sequential DEVS simulations. They found
that Ladder Queue outperformed every other priority queue based
pending event data structure such as Sorted List, Minimal List, Binary
Heap, Splay Tree, and Calendar Queue. Tang~\cite{tang-05} and
Franceschini~\cite{franceschini-15} both use the classic Hold
benchmark simulation model used in this research.

In contrast to earlier work, rather than using a linked list based
implementation, we propose an alternative implementation using
dynamically growing arrays, that is, std::vector from the C(++)
library. Furthermore, we trigger \textit{Bottom} to \textit{Ladder}
re-bucketing only if the \textit{Bottom} has events at different
timestamps to reduce inefficiencies. Our 2-tier Ladder Queue
(\textbf{2tLadderQ}) is a novel enhancement to the Ladder Queue to
enable its efficient use in optimistic parallel simulations.

Dickman~\cite{dickman-13} compare event list data structures that
consisted of Splay Tree, STL Multiset and Ladder Queue. However, the
focus of their paper was in developing a framework for handling
pending event set data structure in shared memory PDES. A central
component of their study was the identification of an appropriate data
structure and design for the shared pending event
set. Gupta~\cite{gupta-14} extended their implementation of Ladder
Queue for shared memory Time Warp based simulation environment, so
that it supports lock-free access to events in the shared pending
event set. The modification involved the use of an unsorted lock-free
queue in the underlying Ladder Queue
structure. Marotta~\cite{marotta-16} contributed to the study of
pending event set data structures in threaded PDES through the design
of the Non-Blocking Priority Queue (NBPQ) data structure. A pending
event set data structure that is closely related to Calendar Queues
with constant time performance.

In contrast to aforementioned efforts, our research focuses on
distributed memory platforms in which each parallel process is single
threaded. Consequently, our implementation does not involve thread
synchronization issues. However, our 2-tier design has the ability to
further reduce lock contention issues in multithreaded environments
and could provide further performance boost. To the best of our
knowledge, the Fibonnacci heap (\textbf{fibHeap}) and our 3-tier Heap
(\textbf{3tHeap}) are unique data structures that have potential to be
effective in simulations with high concurrency.

    

    
